Extreme value analysis for the sample autocovariance matrices of heavy-tailed multivariate time series

### 2017-08-25 papers/DavisMikoschHeinyXie/PaperBody.tex
Probably the final revision of the thesis.

\section{Estimation of the largest eigenvalues: an overview in the iid case}\label{sec:motivation}\setcounter{equation}{0}

@@

It turns out that limit theory for the largest eigenvalues becomes ``easier''
when the dimension $p$ increases with $n$.
Over the last 15 years there has been increasing interest in the case when $p=p_n\to\infty$ as $\nto$. In most of
the literature (exceptions are El Karoui \cite{elkaroui:2003}, Davis et al. \cite{davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014} and Heiny and Mikosch
\cite{heiny:mikosch:2015:iid})
one assumes that $p$ and $n$ grow at the same rate:
\beam\label{eq:gamma}
\dfrac p n\to \gamma\qquad \mbox{for some $\gamma \in (0,\infty)$.}
\eeam
%One explanation for the commonly used condition \eqref{eq:gamma} is that much of the theory was developed for $n \times n$ {\em Wigner matrices}, which is a Hermitean matrix $H$ such that $\{ H_{ij}:i<j \}$ and $\{ H_{ii}:0\le i\le n \}$ are two sets of iid random variables. If $p =\gamma n$, it turns out that the limit theory for the eigenvalues of Wigner matrices and sample covariance matrices is closely connected by the convergence of the empirical spectral distributions to the semicircle and the Mar\v cenko--Pastur law, respectively; see Bai and Silverstein \cite{bai:silverstein:2010}, Chapters 1-3.

@@

Sections~\ref{sec:samplecov} and \ref{sec:possemidef} are devoted to {\em sample autocovariance matrices}.  Motivated by \cite{lam:yao}, we study the eigenvalues of sums of transformed matrices and illustrate the results in two examples. These results are applied to the time series of S\&P 500 in Section \ref{sec:sp500}.
%The technical proofs are collected in Section~\ref{sec:proof}. 
Appendix~\ref{appendix:A} contains useful facts about regular variation and point processes.

\section{General growth rates for $p_n$ in the iid heavy-tailed case}
\label{sec:Davis2}\setcounter{equation}{0}
This section is based on ideas in Heiny and Mikosch \cite{heiny:mikosch:2015:iid} where one can also find detailed proofs.
\subsection*{Growth conditions on $(p_n)$}\label{subsec:pn}
In many applications it is not realistic to assume
that the dimension $p$ of the data and the sample size $n$ grow at the same rate.
The aforementioned results of Soshnikov~\cite{soshnikov:2004,soshnikov:2006} and Auffinger et al. \cite{auffinger:arous:peche:2009} already indicate that the value $\gamma$ in the growth
rate~\eqref{eq:gamma} does not appear in the \ds al limits.
This obervation is in contrast to the light-tailed case; see \eqref{eq:geman}
and \eqref{eq:tc}.  Davis et al.~\cite{davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014}
and Heiny and Mikosch ~\cite{heiny:mikosch:2015:iid} allowed for more general rates for
$p_n\to\infty$ than linear growth in $n$.
Recall that $p=p_n\to\infty$ is the number of rows in the
matrix $\X_n$. We need to specify the growth rate of $(p_n)$ to ensure
a non-degenerate limit distribution of the normalized singular values of the sample autocovariance
matrices. To be precise, we assume
\begin{equation}\label{eq:p}
p=p_n=n^\beta \ell(n), \quad n\ge1,\tag{$C_p(\beta)$}
\end{equation}
where $\ell$ is a slowly varying function and $\beta\ge 0$. If $\beta =0$, we also assume $\ell(n) \to \infty$.
Condition~\ref{eq:p} is more general than the growth conditions in the literature; see
\cite{auffinger:arous:peche:2009,davis:mikosch:pfaffel:2015,davis:pfaffel:stelzer:2014}.
\begin{theorem}\label{thm:intro}
Assume that $\X=\X_n$ has iid entries satisfying the \regvar\ condition \eqref{eq:regvar} for some
$\alpha \in (0,4)$. If $\E[|X|]<\infty$ we also suppose that $\E [X]=0$. Let $(p_n)$ be an integer sequence
satisfying \ref{eq:p} with $\beta\ge 0$. In addition, we require
\begin{equation}\label{Cbeta}
\min(\beta,\beta^{-1})\in (\alpha/2-1,1] \qquad \mbox{ for } \alpha\in [2,4), \tag{$\widetilde{C}_\beta(\alpha)$}
\end{equation}
Then
\beam\label{eq:Davis2}
\sum_{i=1}^p \vep_{a_{np}^{-2}\la_i}\std \sum_{i=1}^\infty \vep_{\Gamma_i^{-2/\alpha}}\,,\qquad \nto\,,
\eeam
where the \con\ holds in the space of point \ms s with state space $(0,\infty)$ equipped with the vague toplogy; see Resnick \cite{resnick:2007}.
\end{theorem}

@@

\par
Proposition~\ref{prop:offdiagonal} is not unexpected for two reasons:
\begin{itemize}
\item
It is well-known from classical theory (see Embrechts and Veraverbeke \cite{embrechts:veraverbeke:1982}) that for any iid \regvary\ non-negative \rv s
$Y,Y'$ with index  $\alpha'>0$, $Y\,Y'$ is \regvary\ with index $\alpha'$ while $Y^2$ is \regvary\ with index $\alpha'/2$.
Therefore $X^2$ and $X_{11}X_{12}$  are \regvary\ with indices $\alpha/2$ and $\alpha$, respectively.
\item
The aforementioned tail behavior is inherited by the entries of $\X\X'$ in the following sense.
By virtue of Nagaev-type \ld\ results for an iid \regvary\ \seq\ $(Y_i)$ with index $\alpha'\in (0,2)$ where we also assume that
$\E [Y_0]=0$ if $\E[|Y_0|]<\infty$
(see Theorem~\ref{thm:nagaev}) %S.V. Nagaev \cite{nagaev:1979} and
%Cline and Hsing \cite{cline:hsing:1998})
we have that $\P(Y_1+\cdots +Y_n >b_n)/(n \,\P(|Y_0|>b_n))$ converges to a non-negative constant
provided $b_n/a_n'\to\infty$, where $\P(|Y_0|>a_n')\sim n^{-1}$ as $\nto$. As a con\seq\ of the tail behaviors of $X_{it}^2$ and $X_{it}X_{jt}$ for $i\ne j$
and Nagaev's results we have for $(b_n)$ \st\ $b_n/a_n^2\to\infty$,
\beao
\dfrac{\P\big((\X\X')_{ij}> b_n\big)}{\P\big((\X\X')_{ii}- c_n> b_n\big)}\sim \dfrac{n\,\P( X_{11}X_{12}>b_n)}{n\,\P(X^2>b_n)}\to 0\,,\qquad \nto\,,
\eeao
where $c_n=0$ or $n\,\E[X^2]$ according as $\alpha\in (0,2)$ or $\alpha\in (2,4)$.
This means that the diagonal and off-diagonal entries of
$\X\X'$ inherit the tails of $X_{it}^2$ and $X_{it}X_{jt}$, $i\ne j$, respectively, above the high threshold $b_n$.
\end{itemize}
\par
Proposition~\ref{prop:offdiagonal} has some immediate con\seq s for the approximation of the eigenvalues
of $\X\X'$ by those of ${\rm diag}(\X\X')$. Indeed, let $C$ be a symmetric $p\times p$ matrix with
eigenvalues $\la_1(C),\ldots,\la_p(C)$ and
ordered eigenvalues
\beam \label{eq:weyl}
\la_{(1)}(C)\ge \cdots \ge \la_{(p)}(C)\,.
\eeam
Then for any symmetric $p\times p$ matrices $A,B$, by {\em Weyl's inequality} (see Bhatia \cite{bhatia:1997}),
\beao
\max_{i=1,\ldots,p}\big|\la_{(i)}(A+B)-\la_{(i)}(A)\big|\le \|B\|_2\,.
\eeao
If we now choose
$A+B=\X\X'$ and $A= \diag (\X\X')$ we obtain the following result.
\begin{corollary}\label{cor:687}
Under the conditions of Proposition~\ref{prop:offdiagonal},
\beao
a_{np}^{-2}\,\max_{i=1,\ldots,p}\big|\la_{(i)}-\la_{(i)}(\diag(\X\X'))\big|\stp 0\,,\quad\nto \,.
\eeao
\end{corollary}
Thus the problem of deriving limit theory for the order statistics of $\X\X'$ has been reduced to limit theory for the order
statistics of the iid row-sums
\beao
D_i^{\rightarrow}= (\X\X')_{ii}=\sum_{t=1}^n X_{it}^2\,,\qquad i=1,\ldots,p\,,
\eeao
which are the eigenvalues of $\diag(\X\X')$. This theory is completely described by the \pp es constructed from the points $D_i^\rightarrow/a_{np}^2$ $i=1,\ldots,p$. Necessary
and sufficient conditions for the weak \con\ of these \pp es are provided by  Lemma~\ref{lem:ppr}
which in combination with the Nagaev-type \ld\ results of  Theorem~\ref{thm:nagaev}
yield the following result; see also Davis et al.~\cite{davis:mikosch:pfaffel:2015}.
\begin{lemma}\label{lem:pp}
Assume the conditions of Proposition~\ref{prop:offdiagonal} hold. Then
\beao
\sum_{i=1}^p \vep_{a_{np}^{-2}(D_i^\rightarrow-c_n)}\std \sum_{i=1}^\infty \vep_{\Gamma_i^{-2/\alpha}}\,,\qquad \nto\,,
\eeao
where $(\Gamma_i)$ is defined in \eqref{eq:Gamma} and $c_n=0$ if $\E[D^\rightarrow]=\infty$ and $c_n=\E[D^\rightarrow]$ otherwise.
\end{lemma}
In this result, centering is only needed for $\alpha\in [2,4)$ when $n/a_{np}^2\not\to 0$. Under the additional condition \ref{Cbeta},
%\begin{equation}\label{Cbetaor}
%\beta\in (\alpha/2-1,1], \quad \mbox{ for } \alpha\in [2,4),
%\end{equation}
$n/a_{np}^2\to 0$ in view of the Potter bounds; see Bingham et al. \cite[p.~25]{bingham:goldie:teugels:1987}.
Combining Lemma~\ref{lem:pp} and Corollary~\ref{cor:687}, we conclude that Theorem~\ref{thm:intro} holds for $\beta\in [0,1]$.

@@

%-----------------------------------------------------------------------
\section{Introducing dependence between the rows and columns}\label{sec:model}\setcounter{equation}{0}
For details on the results of this section, we refer to Davis et al.~\cite{davis:mikosch:pfaffel:2015}, Heiny and Mikosch \cite{heiny:mikosch:2015:iid} and Heiny et al.~\cite{heiny:mikosch:2016:noniid}.

@@

present form by Davis et al. \cite{davis:mikosch:pfaffel:2015}.

@@

originates from a paper by Lam and Yao~\cite{lam:yao} who used a model different from \eqref{eq:Davis1}.

@@

\subsubsection*{The matrix $\M(s)$} We introduce some auxiliary matrices

@@

\H(s)=(h_{k,l+s})_{k,l\in \Z}, \qquad \M(s)= \H(0)\H(s)'\qquad s\ge 0\,.

@@

(\M(s))_{ij}= \sum_{l\in \Z} h_{i,l} h_{j,l+s}, \qquad i,j \in \Z .

@@

We denote the ordered singular values of $\M(s)$ by

@@

Let $r(s)$ be the rank of $\M(s)$ so that $v_{r(s)}(s)>0$ while $v_{r(s)+1}(s)=0$ if $r(s)$ is

@@

\sum_{i=1}^\infty (v_i(s))^2 &=&  \frobnorm{\M(s)}^2= \sum_{i,j\in \Z} \sum_{l_1,l_2 \in \Z} h_{i,l_1} h_{j,l_1+s}h_{i,l_2}

@@

with more restrictive conditions on the growth rate of $(p_n)$ is contained in Davis et al.~\cite{davis:mikosch:pfaffel:2015}.

@@

$(h_{kl})$, hence  unconditionally in $\P$-probability; see also \cite{davis:mikosch:pfaffel:2015}.\end{remark}

@@

Theorem~\ref{thm:mains}  and arguments similar to the proofs in Davis et al. \cite{davis:mikosch:pfaffel:2015}

@@

\section{Some applications}\setcounter{equation}{0}
\subsection{Sample covariance matrices}\label{sec:samplecov}%\setcounter{equation}{0}
The sample covariance matrix $\X_n(0)\X_n(0)'=\X\X'$ is a non-negative definite matrix. Therefore
its eigenvalues and singular values coincide.  Moreover, $v_j=v_j(0)$, $j\ge 1$, are the eigenvalues of $\M=\M(0)$.
\par
Theorem~\ref{thm:mains}(1) yields an approximation of the ordered eigenvalues $(\la_{(i)})$ of $\X\X'$ by
the quantities $(\delta_{(i)})$ which are derived from the order statistics of $(Z_{it}^2)$.
Part (2) of this result provides an approximation of $(\la_{(i)})$ by the quantities
$(\gamma_{(i)}^{\rightarrow/\downarrow})$ which are derived from the order statistics
of the partial sums $(D_i^{\rightarrow/\downarrow})$.
\par
In the following example we illustrate the
quality of the two approximations.
\begin{example}\label{ex:xiao}\em
We choose a Pareto-type \ds\ for $Z$ with density
\beam\label{eq:distrsim}
f_Z(x) =
\left\{\begin{array}{cc}
 \frac{\alpha}{(4|x|)^{\alpha + 1}}\,, & \mbox{if } |x| > 1/4 \\
1\,, & \mbox{otherwise.}
\end{array}\right.
\eeam
We simulated $20,000$ matrices $\X_n$ for $n=1,000$ and $p=200$ whose
iid entries have this density. We assume $\beta=1$.
Note that $\M=\M(0)$ has rank one and $v_1=1$.
The estimated densities of the deviations $a_{np}^{-2}(\lambda_{(1)}-D_{(1)}^\rightarrow)$ and
$a_{np}^{-2}(\lambda_{(1)}-Z^2_{(1),np})$  based on the simulations are shown in
Figure~\ref{fig:lambda_comparison}. The approximation error is very small indeed.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.40]{lambda_comparison.pdf}
  \caption{Density of the approximation errors for the eigenvalues of $a_{np}^{-2}\X\X'$.
The entries of $\X$ are iid with density \eqref{eq:distrsim} and $\alpha=1.6$.}
  \label{fig:lambda_comparison}
\end{figure}
According to the theory,
\beao
a_{np}^{-2}\sup_i|D_{(i)}^\rightarrow-\lambda_{(i)}| +a_{np}^{-2}\sup_i |Z^2_{(i),np}-\lambda_{(i)}|\stp 0\,,
\eeao
but  for finite $n$ the $(D_{(i)}^\rightarrow)$ sequence yields a better approximation to $(\la_{(i)})$. By construction,
the considered differences
have a tendency to be positive but Figure~\ref{fig:lambda_comparison} also shows that the median of the
approximation error for $a_{np}^{-2}(\lambda_{(1)}-D_{(1)}^\rightarrow)$ is almost zero.
\end{example}
\par
Theorem~\ref{cor:1} and the continuous mapping theorem immediately yield results about
the joint convergence of the largest eigenvalues of the matrices $a_{np}^{-2}\X_n\X_n'$ for $\alpha\in (0,2)$ and
$\alpha\in (2,4)$ when $\beta$ satisfies \ref{Cbeta}.
For fixed $k\ge 1$ one gets
\begin{equation*}
a_{np}^{-2} \big(\la_{(1)},\ldots,\la_{(k)}\big)\cid
\big(d_{(1)},\ldots, d_{(k)}\big)\,,
\end{equation*}
where $d_{(1)}\ge \cdots\ge d_{(k)}$ are the $k$ largest ordered
values of the set $\{\Gamma_i^{-2/\alpha} v_j, i=1,2,\ldots,j=1,\ldots,r\}$.
The continuous mapping theorem yields for $k\ge 1$,
\begin{equation}\label{eq:u}
\dfrac{\la_{(1)}}{\la_{(1)}+\cdots+ \la_{(k)}}\cid
\dfrac{d_{(1)}}{d_{(1)}+\cdots+ d_{(k)}}\,,\quad \nto\,.
\end{equation}

An application of the continuous mapping theorem to  the distributional convergence of the point processes in Theorem~\ref{cor:1} in the spirit of Resnick
\cite{resnick:2007}, Theorem 7.1, also yields the following result; see Davis et al. \cite{davis:mikosch:pfaffel:2015} for a proof and a similar result in the case $\alpha \in (2,4)$.
\begin{corollary}\label{cor:1q} Assume the conditions of Theorem~\ref{thm:mains}.
If $\alpha \in (0,2]$ and $\E[Z^2]=\infty$, then
\begin{equation*}
a_{np}^{-2}\Big(\la_{(1)},\sum_{i=1}^p \la_i\Big) \cid
\Big( v_1\,\Gamma_1^{-2/\alpha}\,,\sum_{j=1}^r v_j\,\sum_{i=1}^\infty \Gamma_i^{-2/\alpha}\Big)\,,
\end{equation*}
where $\Gamma_1^{-2/\alpha}$ is Fr\'echet $\Phi_{\alpha/2}$-distributed.
and $\sum_{i=1}^\infty \Gamma_i^{-2/\alpha}$ has the distribution
of a positive $\alpha/2$-stable random variable.
In particular,
\begin{equation}\label{eq:limit}
\dfrac{\la_{(1)}}{\la_{1}+\cdots+ \la_{p}}\cid
\dfrac{v_1}{\sum_{j=1}^r v_j}\;
\dfrac{\Gamma_1^{-2/\alpha}}{\sum_{i=1}^\infty\Gamma_i^{-2/\alpha}}\,,\quad
\nto\,.
\end{equation}
\end{corollary}

@@

when the rank $r$ of the matrix $\M$ is larger than one.

@@

\M= \sum_{l\in \Z} c_l^2\;\; (\theta_i\theta_j)_{i,j \in \Z}\,.

@@

\section{Auxiliary results}\label{appendix:A}\setcounter{equation}{0}

>>>>

%% \section{Estimation of the largest eigenvalues: an overview in the iid case}\label{sec:motivation}
\section{Estimation of the largest eigenvalues}
\label{sec:motivation}
% \setcounter{equation}{0}

@@

It turns out that limit theory for the largest eigenvalues becomes ``easier''
when the dimension $p$ increases with $n$.
Over the last 15 years there has been increasing interest in the case when $p=p_n\to\infty$ as $\nto$. In most of
the literature (exceptions are El Karoui \cite{elkaroui:2003}, Davis et al. \cite{davis:mikosch:pfaffel:2016,davis:pfaffel:stelzer:2014} and Heiny and Mikosch
\cite{heiny:mikosch:2016})
one assumes that $p$ and $n$ grow at the same rate:
\beam\label{eq:gamma}
\dfrac p n\to \gamma\qquad \mbox{for some $\gamma \in (0,\infty)$.}
\eeam
%One explanation for the commonly used condition \eqref{eq:gamma} is that much of the theory was developed for $n \times n$ {\em Wigner matrices}, which is a Hermitean matrix $H$ such that $\{ H_{ij}:i<j \}$ and $\{ H_{ii}:0\le i\le n \}$ are two sets of iid random variables. If $p =\gamma n$, it turns out that the limit theory for the eigenvalues of Wigner matrices and sample covariance matrices is closely connected by the convergence of the empirical spectral distributions to the semicircle and the Mar\v cenko--Pastur law, respectively; see Bai and Silverstein \cite{bai:silverstein:2010}, Chapters 1-3.

@@

Sections~\ref{sec:samplecov} and \ref{sec:possemidef} are devoted to {\em sample autocovariance matrices}.  Motivated by \cite{lam:yao:2012}, we study the eigenvalues of sums of transformed matrices and illustrate the results in two examples. These results are applied to the time series of S\&P 500 in Section \ref{sec:sp500}.
%The technical proofs are collected in Section~\ref{sec:proof}. 
Appendix~\ref{appendix:A} contains useful facts about regular variation and point processes.

\section{General growth rates for $p_n$ in the iid heavy-tailed case}
\label{sec:Davis2}
% \setcounter{equation}{0}
This section is based on ideas in Heiny and Mikosch \cite{heiny:mikosch:2016} where one can also find detailed proofs.
\subsection*{Growth conditions on $(p_n)$}\label{subsec:pn}
In many applications it is not realistic to assume
that the dimension $p$ of the data and the sample size $n$ grow at the same rate.
The aforementioned results of Soshnikov~\cite{soshnikov:2004,soshnikov:2006} and Auffinger et al. \cite{auffinger:arous:peche:2009} already indicate that the value $\gamma$ in the growth
rate~\eqref{eq:gamma} does not appear in the \ds al limits.
This obervation is in contrast to the light-tailed case; see \eqref{eq:geman}
and \eqref{eq:tc}.  Davis et al.~\cite{davis:mikosch:pfaffel:2016,davis:pfaffel:stelzer:2014}
and Heiny and Mikosch ~\cite{heiny:mikosch:2016} allowed for more general rates for
$p_n\to\infty$ than linear growth in $n$.
Recall that $p=p_n\to\infty$ is the number of rows in the
matrix $\X_n$. We need to specify the growth rate of $(p_n)$ to ensure
a non-degenerate limit distribution of the normalized singular values of the sample autocovariance
matrices. To be precise, we assume
\begin{equation}\label{eq:p}
p=p_n=n^\beta \ell(n), \quad n\ge1,\tag{$C_p(\beta)$}
\end{equation}
where $\ell$ is a slowly varying function and $\beta\ge 0$. If $\beta =0$, we also assume $\ell(n) \to \infty$.
Condition~\ref{eq:p} is more general than the growth conditions in the literature; see
\cite{auffinger:arous:peche:2009,davis:mikosch:pfaffel:2016,davis:pfaffel:stelzer:2014}.
\begin{theorem}\label{thm:intro}
Assume that $\X=\X_n$ has iid entries satisfying the \regvar\ condition \eqref{eq:regvar} for some
$\alpha \in (0,4)$. If $\E[|X|]<\infty$ we also suppose that $\E [X]=0$. Let $(p_n)$ be an integer sequence
satisfying \ref{eq:p} with $\beta\ge 0$. In addition, we require
\begin{equation}\label{Cbeta}
\min(\beta,\beta^{-1})\in (\alpha/2-1,1] \qquad \mbox{ for } \alpha\in [2,4), \tag{$\widetilde{C}_\beta(\alpha)$}
\end{equation}
Then
\beam\label{eq:Davis2}
\sum_{i=1}^p \vep_{a_{np}^{-2}\la_i}\std \sum_{i=1}^\infty \vep_{\Gamma_i^{-2/\alpha}}\,,\qquad \nto\,,
\eeam
where the \con\ holds in the space of point \ms s with state space $(0,\infty)$ equipped with the vague toplogy; see Resnick \cite{resnick:2007}.
\end{theorem}

@@

\par
Proposition~\ref{prop:offdiagonal} is not unexpected for two reasons:
\begin{itemize}
\item
It is well-known from classical theory (see Embrechts and Veraverbeke \cite{embrechts:veraverbeke:1982}) that for any iid \regvary\ non-negative \rv s
$Y,Y'$ with index  $\alpha'>0$, $Y\,Y'$ is \regvary\ with index $\alpha'$ while $Y^2$ is \regvary\ with index $\alpha'/2$.
Therefore $X^2$ and $X_{11}X_{12}$  are \regvary\ with indices $\alpha/2$ and $\alpha$, respectively.
\item
The aforementioned tail behavior is inherited by the entries of $\X\X'$ in the following sense.
By virtue of Nagaev-type \ld\ results for an iid \regvary\ \seq\ $(Y_i)$ with index $\alpha'\in (0,2)$ where we also assume that
$\E [Y_0]=0$ if $\E[|Y_0|]<\infty$
(see Theorem~\ref{thm:nagaev}) %S.V. Nagaev \cite{nagaev:1979} and
%Cline and Hsing \cite{cline:hsing:1998})
we have that $\P(Y_1+\cdots +Y_n >b_n)/(n \,\P(|Y_0|>b_n))$ converges to a non-negative constant
provided $b_n/a_n'\to\infty$, where $\P(|Y_0|>a_n')\sim n^{-1}$ as $\nto$. As a con\seq\ of the tail behaviors of $X_{it}^2$ and $X_{it}X_{jt}$ for $i\ne j$
and Nagaev's results we have for $(b_n)$ \st\ $b_n/a_n^2\to\infty$,
\beao
\dfrac{\P\big((\X\X')_{ij}> b_n\big)}{\P\big((\X\X')_{ii}- c_n> b_n\big)}\sim \dfrac{n\,\P( X_{11}X_{12}>b_n)}{n\,\P(X^2>b_n)}\to 0\,,\qquad \nto\,,
\eeao
where $c_n=0$ or $n\,\E[X^2]$ according as $\alpha\in (0,2)$ or $\alpha\in (2,4)$.
This means that the diagonal and off-diagonal entries of
$\X\X'$ inherit the tails of $X_{it}^2$ and $X_{it}X_{jt}$, $i\ne j$, respectively, above the high threshold $b_n$.
\end{itemize}
\par
Proposition~\ref{prop:offdiagonal} has some immediate con\seq s for the approximation of the eigenvalues
of $\X\X'$ by those of ${\rm diag}(\X\X')$. Indeed, let $C$ be a symmetric $p\times p$ matrix with
eigenvalues $\la_1(C),\ldots,\la_p(C)$ and
ordered eigenvalues
\beam \label{eq:weyl}
\la_{(1)}(C)\ge \cdots \ge \la_{(p)}(C)\,.
\eeam
Then for any symmetric $p\times p$ matrices $A,B$, by {\em Weyl's inequality} (see Bhatia \cite{bhatia:1997}),
\beao
\max_{i=1,\ldots,p}\big|\la_{(i)}(A+B)-\la_{(i)}(A)\big|\le \|B\|_2\,.
\eeao
If we now choose
$A+B=\X\X'$ and $A= \diag (\X\X')$ we obtain the following result.
\begin{corollary}\label{cor:687}
Under the conditions of Proposition~\ref{prop:offdiagonal},
\beao
a_{np}^{-2}\,\max_{i=1,\ldots,p}\big|\la_{(i)}-\la_{(i)}(\diag(\X\X'))\big|\stp 0\,,\quad\nto \,.
\eeao
\end{corollary}
Thus the problem of deriving limit theory for the order statistics of $\X\X'$ has been reduced to limit theory for the order
statistics of the iid row-sums
\beao
D_i^{\rightarrow}= (\X\X')_{ii}=\sum_{t=1}^n X_{it}^2\,,\qquad i=1,\ldots,p\,,
\eeao
which are the eigenvalues of $\diag(\X\X')$. This theory is completely described by the \pp es constructed from the points $D_i^\rightarrow/a_{np}^2$ $i=1,\ldots,p$. Necessary
and sufficient conditions for the weak \con\ of these \pp es are provided by  Lemma~\ref{lem:ppr}
which in combination with the Nagaev-type \ld\ results of  Theorem~\ref{thm:nagaev}
yield the following result; see also Davis et al.~\cite{davis:mikosch:pfaffel:2016}.
\begin{lemma}\label{lem:pp}
Assume the conditions of Proposition~\ref{prop:offdiagonal} hold. Then
\beao
\sum_{i=1}^p \vep_{a_{np}^{-2}(D_i^\rightarrow-c_n)}\std \sum_{i=1}^\infty \vep_{\Gamma_i^{-2/\alpha}}\,,\qquad \nto\,,
\eeao
where $(\Gamma_i)$ is defined in \eqref{eq:Gamma} and $c_n=0$ if $\E[D^\rightarrow]=\infty$ and $c_n=\E[D^\rightarrow]$ otherwise.
\end{lemma}
In this result, centering is only needed for $\alpha\in [2,4)$ when $n/a_{np}^2\not\to 0$. Under the additional condition \ref{Cbeta},
%\begin{equation}\label{Cbetaor}
%\beta\in (\alpha/2-1,1], \quad \mbox{ for } \alpha\in [2,4),
%\end{equation}
$n/a_{np}^2\to 0$ in view of the Potter bounds; see Bingham et al. \cite[p.~25]{bingham:goldie:teugels:1987}.
Combining Lemma~\ref{lem:pp} and Corollary~\ref{cor:687}, we conclude that Theorem~\ref{thm:intro} holds for $\beta\in [0,1]$.

@@

%-----------------------------------------------------------------------
\section{Introducing dependence between the rows and columns}\label{sec:model}
% \setcounter{equation}{0}
For details on the results of this section, we refer to Davis et al.~\cite{davis:mikosch:pfaffel:2016}, Heiny and Mikosch \cite{heiny:mikosch:2016} and Heiny et al.~\cite{heiny:mikosch:2016:noniid}.

@@

present form by Davis et al. \cite{davis:mikosch:pfaffel:2016}.

@@

originates from a paper by Lam and Yao~\cite{lam:yao:2012} who used a model different from \eqref{eq:Davis1}.

@@

\subsubsection*{The matrix $\mathbf M(s)$} We introduce some auxiliary matrices

@@

\H(s)=(h_{k,l+s})_{k,l\in \Z}, \qquad \mathbf M(s)= \H(0)\H(s)'\qquad s\ge 0\,.

@@

(\mathbf M(s))_{ij}= \sum_{l\in \Z} h_{i,l} h_{j,l+s}, \qquad i,j \in \Z .

@@

We denote the ordered singular values of $\mathbf M(s)$ by

@@

Let $r(s)$ be the rank of $\mathbf M(s)$ so that $v_{r(s)}(s)>0$ while $v_{r(s)+1}(s)=0$ if $r(s)$ is

@@

\sum_{i=1}^\infty (v_i(s))^2 &=&  \frobnorm{\mathbf M(s)}^2= \sum_{i,j\in \Z} \sum_{l_1,l_2 \in \Z} h_{i,l_1} h_{j,l_1+s}h_{i,l_2}

@@

with more restrictive conditions on the growth rate of $(p_n)$ is contained in Davis et al.~\cite{davis:mikosch:pfaffel:2016}.

@@

$(h_{kl})$, hence  unconditionally in $\P$-probability; see also \cite{davis:mikosch:pfaffel:2016}.\end{remark}

@@

Theorem~\ref{thm:mains}  and arguments similar to the proofs in Davis et al. \cite{davis:mikosch:pfaffel:2016}

@@

\section{Some applications}
% \setcounter{equation}{0}
\subsection{Sample covariance matrices}\label{sec:samplecov}
The sample covariance matrix $\X_n(0)\X_n(0)'=\X\X'$ is a non-negative definite matrix. Therefore
its eigenvalues and singular values coincide.  Moreover, $v_j=v_j(0)$, $j\ge 1$, are the eigenvalues of $\mathbf M=\mathbf M(0)$.
\par
Theorem~\ref{thm:mains}(1) yields an approximation of the ordered eigenvalues $(\la_{(i)})$ of $\X\X'$ by
the quantities $(\delta_{(i)})$ which are derived from the order statistics of $(Z_{it}^2)$.
Part (2) of this result provides an approximation of $(\la_{(i)})$ by the quantities
$(\gamma_{(i)}^{\rightarrow/\downarrow})$ which are derived from the order statistics
of the partial sums $(D_i^{\rightarrow/\downarrow})$.
\par
In the following example we illustrate the
quality of the two approximations.
\begin{example}\label{ex:xiao}\em
We choose a Pareto-type \ds\ for $Z$ with density
\beam\label{eq:distrsim}
f_Z(x) =
\left\{\begin{array}{cc}
 \frac{\alpha}{(4|x|)^{\alpha + 1}}\,, & \mbox{if } |x| > 1/4 \\
1\,, & \mbox{otherwise.}
\end{array}\right.
\eeam
We simulated $20,000$ matrices $\X_n$ for $n=1,000$ and $p=200$ whose
iid entries have this density. We assume $\beta=1$.
Note that $\mathbf M=\mathbf M(0)$ has rank one and $v_1=1$.
The estimated densities of the deviations $a_{np}^{-2}(\lambda_{(1)}-D_{(1)}^\rightarrow)$ and
$a_{np}^{-2}(\lambda_{(1)}-Z^2_{(1),np})$  based on the simulations are shown in
Figure~\ref{fig:lambda_comparison}. The approximation error is very small indeed.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.40]{lambda_comparison.pdf}
  \caption{Density of the approximation errors for the eigenvalues of $a_{np}^{-2}\X\X'$.
The entries of $\X$ are iid with density \eqref{eq:distrsim} and $\alpha=1.6$.}
  \label{fig:lambda_comparison}
\end{figure}
According to the theory,
\beao
a_{np}^{-2}\sup_i|D_{(i)}^\rightarrow-\lambda_{(i)}| +a_{np}^{-2}\sup_i |Z^2_{(i),np}-\lambda_{(i)}|\stp 0\,,
\eeao
but  for finite $n$ the $(D_{(i)}^\rightarrow)$ sequence yields a better approximation to $(\la_{(i)})$. By construction,
the considered differences
have a tendency to be positive but Figure~\ref{fig:lambda_comparison} also shows that the median of the
approximation error for $a_{np}^{-2}(\lambda_{(1)}-D_{(1)}^\rightarrow)$ is almost zero.
\end{example}
\par
Theorem~\ref{cor:1} and the continuous mapping theorem immediately yield results about
the joint convergence of the largest eigenvalues of the matrices $a_{np}^{-2}\X_n\X_n'$ for $\alpha\in (0,2)$ and
$\alpha\in (2,4)$ when $\beta$ satisfies \ref{Cbeta}.
For fixed $k\ge 1$ one gets
\begin{equation*}
a_{np}^{-2} \big(\la_{(1)},\ldots,\la_{(k)}\big)\cid
\big(d_{(1)},\ldots, d_{(k)}\big)\,,
\end{equation*}
where $d_{(1)}\ge \cdots\ge d_{(k)}$ are the $k$ largest ordered
values of the set $\{\Gamma_i^{-2/\alpha} v_j, i=1,2,\ldots,j=1,\ldots,r\}$.
The continuous mapping theorem yields for $k\ge 1$,
\begin{equation}\label{eq:u}
\dfrac{\la_{(1)}}{\la_{(1)}+\cdots+ \la_{(k)}}\cid
\dfrac{d_{(1)}}{d_{(1)}+\cdots+ d_{(k)}}\,,\quad \nto\,.
\end{equation}

An application of the continuous mapping theorem to  the distributional convergence of the point processes in Theorem~\ref{cor:1} in the spirit of Resnick
\cite{resnick:2007}, Theorem 7.1, also yields the following result; see Davis et al. \cite{davis:mikosch:pfaffel:2016} for a proof and a similar result in the case $\alpha \in (2,4)$.
\begin{corollary}\label{cor:1q} Assume the conditions of Theorem~\ref{thm:mains}.
If $\alpha \in (0,2]$ and $\E[Z^2]=\infty$, then
\begin{equation*}
a_{np}^{-2}\Big(\la_{(1)},\sum_{i=1}^p \la_i\Big) \cid
\Big( v_1\,\Gamma_1^{-2/\alpha}\,,\sum_{j=1}^r v_j\,\sum_{i=1}^\infty \Gamma_i^{-2/\alpha}\Big)\,,
\end{equation*}
where $\Gamma_1^{-2/\alpha}$ is Fr\'echet $\Phi_{\alpha/2}$-distributed.
and $\sum_{i=1}^\infty \Gamma_i^{-2/\alpha}$ has the distribution
of a positive $\alpha/2$-stable random variable.
In particular,
\begin{equation}\label{eq:limit}
\dfrac{\la_{(1)}}{\la_{1}+\cdots+ \la_{p}}\cid
\dfrac{v_1}{\sum_{j=1}^r v_j}\;
\dfrac{\Gamma_1^{-2/\alpha}}{\sum_{i=1}^\infty\Gamma_i^{-2/\alpha}}\,,\quad
\nto\,.
\end{equation}
\end{corollary}

@@

when the rank $r$ of the matrix $\mathbf M$ is larger than one.

@@

\mathbf M= \sum_{l\in \Z} c_l^2\;\; (\theta_i\theta_j)_{i,j \in \Z}\,.

@@

\section{Auxiliary results}\label{appendix:A}
% \setcounter{equation}{0}
